Title Goes Here

### 2023-02-07 inst/paper/_results.qmd
clean up code

```{r}
#| label: detection-tbl-setup

uhplc_calculus_detect <- uhplc_calculus_long %>%
#uhplc_calculus_detect <- uhplc_calculus_demography %>%
  #remove_missing() %>%
  ungroup() %>% 
  filter(quant > 0) %>%
  distinct(compound, batch) %>%
  pivot_wider(names_from = batch, values_from = batch) %>% 
  mutate(across(!compound, ~ if_else(!is.na(.x), TRUE, FALSE)))
```

@@

```{r}
#| label: fig-detection-preservation
#| fig-cap: "(A) Violin plot with overlaid box plots depicting the distribution of extracted quantities of each compound from batch 2 separated by state of preservation of the skeleton. (B) Extracted quantity (ng) of compound plotted against weights of the calculus samples from batch 2."

preservation_plot <- uhplc_calculus_long %>%
  filter(
    !is.na(preservation),
    quant > 0,
    batch == "batch2",
    #!compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = preservation, y = quant)) +
    geom_violin(aes(fill = preservation), alpha = 0.6) +
    #geom_jitter(width = 0.2) +
    geom_boxplot(width = 0.2) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v", ncol = 1) +
    theme_bw() +
    theme(
      legend.position = "none",
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
      ) +
    labs(x = "Preservation", y = "Quantity (ng)") +
    scale_fill_viridis_d()

weight_quant_plot <- uhplc_calculus_long %>%
  filter(
    #!is.na(preservation),
    batch == "batch2",
    quant > 0,
    #!compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = weight, y = quant)) +
    geom_point(col = viridisLite::viridis(1)) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v", ncol = 1) +
    theme_bw() +
    theme(axis.title.y = element_blank()) +
    labs(y = "Quantity (ng)", x = "Weight (mg)")

# weight_quant_plot <- uhplc_data_comb %>%
#   select(sample, contains("batch2")) %>% 
#   select(sample, batch2_weight, contains("calc")) %>% 
#   pivot_longer(
#     -c(sample, batch2_weight),
#     names_to = c("compound", "batch"),
#     names_pattern = "(.*)_(.*)",
#     values_to = c("conc")
#   ) %>% 
#   mutate(compound = str_remove(compound, "_calc")) %>%
#   #remove_missing() %>% 
#   filter(
#     conc > 0,
#     id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
#     ) %>% 
#   ggplot(aes(x = batch2_weight, y = conc)) +
#     geom_point(col = viridisLite::viridis(1)) +
#     facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v") +
#     theme_bw() +
#     theme(axis.title.y = element_blank()) +
#     labs(y = "Quantity (ng)", x = "Weight (mg)")

@@

tobacco_accuracy <- tobacco %>%
  remove_missing(vars = "quant") %>%
  group_by(sample, id, .drop = F) %>% 
  summarise(
    detection = sum(detection),
    #compound = compound
    ) %>% 
  left_join(select(demography, id, pipe_notch, preservation, age), by = "id") %>% 
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(
      detection == 0 & pipe_notch == "N" ~ 1,
      detection > 0 & pipe_notch == "Y" ~ 1,
      detection > 0 & pipe_notch == "N" ~ NaN, # no way of knowing macroscopically if the person smoked without pipe
      TRUE ~ 0
    )
  )

@@

corr_tib <- conc_cor %>% 
  as_tibble(rownames = "var") %>% 
  filter(var != "age",
         var != "periodont_status") #%>% 
  #mutate(across(everything(), signif, 3))

corr_tib_long <- corr_tib %>% 
  pivot_longer(
    -var,
    names_to = "name",
    values_to = "corr"
  )
# still need to remove everything under the diagonal (colour white?)
#corr_tib[corr_tib == 1] <- kableExtra::cell_spec(corr_tib[corr_tib == 1], color = "transparent")
corr_tib_out <- corr_tib %>%
  column_to_rownames("var") %>%
  as.matrix() %>% 
  signif(digits = 2)
cols <- 0
for(i in 7:nrow(corr_tib)){ # terrible solution for formatting table
  cols <- cols + 1 
  corr_tib_out[i,1:cols] <- ""
} 
corr_names <- c("Caries", "Nicotine", "SA", "Calculus", "PN", "Theophylline", "Caffeine", "Cotinine")
row.names(corr_tib_out) <- c(row.names(corr_tib_out)[1:6], corr_names)
corr_tib_out <- corr_tib_out[-nrow(corr_tib),] # remove last row because it has no values (all duplicate coefficients)
corr_tib_out %>% 
  knitr::kable(col.names = corr_names)
```

>>>>

```{r}
#| label: detection-tbl-setup
uhplc_calculus_detect <- uhplc_calculus_long %>%
#uhplc_calculus_detect <- uhplc_calculus_demography %>%
  #remove_missing() %>%
  ungroup() %>% 
  filter(quant > 0) %>%
  distinct(compound, batch) %>%
  pivot_wider(names_from = batch, values_from = batch) %>% 
  mutate(across(!compound, ~ if_else(!is.na(.x), TRUE, FALSE)))
```

@@

```{r}
#| label: fig-detection-preservation
#| fig-cap: "(A) Violin plot with overlaid box plots depicting the distribution of extracted quantities of each compound from batch 2 separated by state of preservation of the skeleton. (B) Extracted quantity (ng) of compound plotted against weights of the calculus samples from batch 2."
#| fig-width: 5
#| fig-height: 7

preservation_plot <- uhplc_calculus_long %>%
  filter(
    !is.na(preservation),
    quant > 0,
    batch == "batch2",
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = preservation, y = quant)) +
    geom_violin(aes(fill = preservation), alpha = 0.6) +
    #geom_jitter(width = 0.2) +
    geom_boxplot(width = 0.2) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v", ncol = 1) +
    theme_bw() +
    theme(
      legend.position = "none",
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
      ) +
    labs(x = "Preservation", y = "Quantity (ng)") +
    scale_fill_viridis_d()

weight_quant_plot <- uhplc_calculus_long %>%
  filter(
    batch == "batch2",
    quant > 0,
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = weight, y = quant)) +
    geom_point(col = viridisLite::viridis(1)) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v", ncol = 1) +
    theme_bw() +
    theme(axis.title.y = element_blank()) +
    labs(y = "Quantity (ng)", x = "Weight (mg)")

@@

tobacco_accuracy <- tobacco %>%
  remove_missing(vars = "quant") %>%
  group_by(sample, id, .drop = F) %>% 
  summarise(
    detection = sum(detection),
    ) %>% 
  left_join(select(demography, id, pipe_notch, preservation, age), by = "id") %>% 
  mutate(
    pipe_notch = if_else(pipe_notch > 0, "Y", "N"),
    correct = case_when(
      detection == 0 & pipe_notch == "N" ~ 1,
      detection > 0 & pipe_notch == "Y" ~ 1,
      detection > 0 & pipe_notch == "N" ~ NaN, # no way of knowing macroscopically if the person smoked without pipe
      TRUE ~ 0
    )
  )

@@

corr_tib <- conc_cor %>% 
  as_tibble(rownames = "var") %>% 
  filter(var != "age",
         var != "periodont_status")

corr_tib_long <- corr_tib %>% 
  pivot_longer(
    -var,
    names_to = "name",
    values_to = "corr"
  )

corr_tib_out <- corr_tib %>%
  column_to_rownames("var") %>%
  as.matrix() %>% 
  signif(digits = 2)
cols <- 0
for(i in 7:nrow(corr_tib)){ # remove repeated values
  cols <- cols + 1 
  corr_tib_out[i,1:cols] <- ""
} 
corr_names <- c("Caries", "Nicotine", "SA", "Calculus", "PN", "Theophylline", "Caffeine", "Cotinine")
row.names(corr_tib_out) <- c(row.names(corr_tib_out)[1:6], corr_names)
corr_tib_out <- corr_tib_out[-nrow(corr_tib),] # remove last row because it has no values (all duplicate coefficients)
corr_tib_out %>% 
  knitr::kable(col.names = corr_names)
```
